A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-17 tex/paper.tex
sean's intermediate edits

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M. Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
%\email{iczekala@cfa.harvard.edu}

@@

A single stellar spectrum contains a vast amount of information about the physical properties of the star. It is of supreme interest to extract a maximal amount of this information from the spectrum in order to study the fundamental stellar parameters such as effective temperature, surface gravity, and metallicity. Accurate and unbiased estimates of stellar parameters are crucial to the study of numerous fields of astrophysics. For example, the mass, radius, and temperature of every newly discovered exoplanet (save for directly imaged planets) are measured relative to the properties of its host star. Large uncertainties or biases in the stellar parameters will translate into incorrect estimates of the habitable zone and the occurrence rate of earth-like planets \citep{kane14}. High quality stellar parameters are also necessary for the study of stellar evolution. Modelers of stellar evolution rely upon accurate samples of stellar parameters to chart the evolutionary tracks of stars through the Hertsprung-Russell diagram.

However, extracting physical parameters of stars from a stellar spectrum is often difficult, due to several factors. First, generating synthetic stellar spectra to use as a benchmark for comparison requires a sophisticated model of the stellar atmosphere and radiative transfer, along with an extensive, high-quality database of opacity sources. Accurately calibrating the atomic constants in this database is a laborious task, and requires high resolution, high signal-to noise spectra for verification, which limits targets mainly to extremely bright stars like the Sun and Vega. This makes it difficult to achieve the same precision when extending synthetic stellar models to intrinsically fainter stars, like M dwarfs, which have additional sources of molecular opacity. Second,  synthesizing a spectrum with such a sophisticated model requires a large amount of computational power, such that a full-optical spectrum at high resolution will take a day or more to process on a supercomputer.

Several, well-tested techniques exist that use spectra to infer stellar parameters. The most widely used technique relies upon measuring ratios of lines, or line indexes, which are matched to a set of indices measured from a synthetic spectrum. More sophisticated codes exist that synthesize spectra for direct comparison with certain well-studied lines, (e.g., \ion{Fe}{0}, \ion{Mg}{0}, and \ion{Ca}{0}) such as MOOG \citep{sneden73} and SME \citep{valenti96}. Other codes such as SPC \citep{buchhave12} cross-correlate the observed spectrum against a grid of pre-computed synthetic spectra and then fit a surface to the normalized cross-correlation coefficients to determine parameter uncertainties.

Owing to the high quality of libraries of pre-computed synthetic stellar spectra, it is now possible to generate an accurate forward model of a stellar spectrum. We seek to place this model in a Bayesian framework and construct a likelihood function suitable for comparing synthetic spectral models to real spectral data. We use a non-trivial covariance matrix, parameterized by a Gaussian process covariance kernel, to account for the residual pixel correlations that arise from a spectroscopic fit, which is a more complex problem than can be addressed with a simple $\chi^2$ best-fit metric. While designed for fitting stellar spectra, our method is general for all comparisons of spectra, and could be used to fit any type of spectra with a synthetic model, (e.g., galaxies or unresolved stellar clusters). In \S\ref{sec:method}, we describe the methodology, including a treatment of the covariance introduced by a spectroscopic fit. In \S\ref{sec:examples}, we present applications of our method to two examples, using optical spectra of an F type star and infrared spectra of an M star. In \S\ref{sec:discussion} we discuss the implications of our method and potential applications to develop data-driven models of stellar spectra.

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help 
characterize the fundamental properties of the observing target.  However, as noted in Section 
\ref{sec:intro}, the reliable and statistically robust inference of those physical (or empirical) 
parameters from an observed spectrum can be extraordinarily challenging.  Here we describe a 
generative Bayesian modeling framework that confronts some of the key obstacles in that process.  
The goal of this approach is to conservatively extract the maximal amount of information about a 
prescribed (and degenerate) parameter set by forward-modeling an observed spectrum, while also 
recognizing and explicitly accounting for the covariances and biases introduced by pathologically 
imperfect models or calibrations.  The method is modular, and therefore can easily incorporate 
additional physical or nuisance parameters as desired without sacrificing an accurate reflection of 
the limitations in the data.  Moreover, with a well-crafted observational sample, this data-driven 
approach should ultimately enable us to systematically learn how synthetic spectral models can be 
improved.  The specific applications discussed here are related to the spectra of individual stars, 
but the methodology is generic (and could be used for composite spectra of unresolved stellar 
clusters, galaxies, etc.).

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{valenti96}.  A more 
common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
technically the former approach is most similar to the ideal case, the computational cost of 
repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
(particularly for data with a large spectral range) considerably less appealing.  A related, but 
different approach is to eschew forward modeling entirely (and therefore repeated spectral 
syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
points of the library.  Then, these discretized samples of the posterior probability density can be 
interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
that this interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
to a typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
n}$, designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

\begin{figure}[!htb]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite. Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000. Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$. Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
\label{fig:broadening}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming a 
derived extinction law $A_{\lambda}$ \citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  
The parameters $\vt_{{\rm obs}, f} = [\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-order (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\vt, \vT_{\rm cheb}) \equiv \vD-\vM(\vt, \vT_{\rm cheb}).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:prob}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnprob}
\end{equation}

>>>>

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
%\email{iczekala@cfa.harvard.edu}

@@

All astronomers recognize that spectroscopy offers a wealth of information that can help
characterize the properties of the observing target.  In the context of stellar astrophysics, 
spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption 
lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface 
gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the 
masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and 
molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]), 
facilitating study of the chemical hallmarks of different stellar populations.  With sufficient 
spectral resolution, the velocity content of a spectrum can convey crucial information about 
stellar rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through 
the radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of 
these stellar properties, there is acute interest in the exoplanet community.  There, all estimates 
of the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host 
mass or radius {\it ratio} is measured with the radial velocity or transit techniques, 
respectively).  Moreover, essential clues to the planet formation process are encapsulated in the 
dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
\citep[e.g.,][]{fischer05,buchhave14}.   
% SA: we ought to check with exoplanet people on these references. Can do later...

That said, the robust and quantitative extraction of physical (or empirical) parameters from an 
observed spectrum can be an extraordinary challenge.  Stellar models are ultimately required as a 
comparative benchmark to associate observed spectral features with the parameters of interest.  
Generating a synthetic model spectrum requires a complex numerical treatment of the stellar 
structure and radiative transfer through the atmosphere \citep[e.g.,][]{kurucz93,castelli04,
hauschildt99,husser13,paxton11}.  Detailed models calibrated to individual stars are important, but 
rare (e.g., the Sun, Vega); as such, these stellar models are relatively untested in large swaths 
of parameter-space.  Moreover, they necessarily include simplifications to treat complicated 
physical processes (e.g., convection) or computational limitations (e.g., geometry, boundary 
conditions), and often must rely on incomplete or inaccurate atomic and molecular information 
(e.g., oscillator strengths, opacities).  In principle, the models could be further improved with 
appropriate reference to spectroscopic datasets.  Nevertheless, they achieve a remarkable level of 
success in reproducing many key diagnostic features in stellar spectra.

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the spectral data, usually in the form of spectral line equivalent widths and/or local 
continuum shapes.  A combined sequence of the ratios of these quantities are found to be especially 
sensitive to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  The line index comparison approach has 
the major advantage of being trivially fast, but each index relationship is only informative over 
a limited range of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some preferential weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained while taking advantage of 
more data content (particularly in the spectral dimension), thereby achieving high precision even 
for data with comparatively low sensitivity; the disadvantage is that the model quality and 
parameter inferences are assessed in an empirical, rather than probabilistic, framework.  The third 
approach of interest prefers a direct, pixel-by-pixel comparison between model and observed 
spectra, and is usually associated with a spectral synthesis front-end (e.g., {\tt MOOG}; 
\citealt{sneden73}; {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric 
flexibility (e.g., one can fit for atomic abundances or arbitrary structures) and a proper 
inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian 
format;][]{shkedy07,schoenrich13}, but it is computationally expensive and can be biased by model 
systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a 
pre-selected subsample of the data.

In this article, we design a flexible forward-modeling approach to the general spectroscopic 
inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
residual pixel-by-pixel correlations that arise when comparing observed spectra to intrinsically 
imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
observations to guide substantially improvements in the models themselves.  An overview of the 
methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
matrix formalism.  Some relevant tests and example applications (specifically for a high resolution 
optical spectrum of an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of 
the method are described in Section \ref{sec:examples}.  Finally, a discussion of the potential 
utility of the technique, and the possibility of extending it to develop data-driven spectral 
models, is provided in Section \ref{sec:discussion}.

@@

Here we describe a generative Bayesian modeling framework that confronts some of the key obstacles 
in the spectroscopic inference problem.  The goal of this approach is to conservatively extract the 
maximal amount of information about a prescribed (and degenerate) parameter set by forward-modeling 
an observed spectrum, while also recognizing and explicitly accounting for the covariances and 
biases introduced by pathologically imperfect models or calibrations.  The method is modular, and 
therefore can easily incorporate additional physical or nuisance parameters as desired without 
sacrificing an accurate reflection of the limitations in the data.  Moreover, with a well-crafted 
observational sample, this data-driven approach should ultimately enable us to systematically learn 
how synthetic spectral models can be improved.  The specific applications discussed here are 
related to the spectra of individual stars, but the methodology is generic (and could be used for 
composite spectra of unresolved stellar clusters, galaxies, etc.).

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for {\tt 
SME};][]{valenti96}.  A more common variant is to instead rely on interpolation over a library of 
pre-synthesized model spectra, $f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,
allard12,husser13}.  While technically the former approach is most similar to the ideal case, the 
computational cost of repeated spectral synthesis is sufficiently high to make a detailed 
exploration of parameter space (particularly for data with a large spectral range) considerably 
less appealing.  A related, but different approach is to eschew forward modeling entirely (and 
therefore repeated spectral syntheses and/or library interpolations), and instead evaluate the 
models only at the discrete grid points of the library.  Then, these discretized samples of the 
posterior probability density can be interpolated to an arbitrary $\vt_{\ast}$ to construct 
appropriate confidence intervals \citep[similar to the method of {\tt SPC};][]{buchhave12}.  The 
difficulty with this latter approach is that the parameter uncertainties can be smaller than the 
grid spacing; in that case, there is valid concern that this interpolation might not accurately 
recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
\comm{I see nothing was done here regarding your updated approach; how should we handle that?}

@@

Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ is highly over-sampled compared to a 
typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
n}$, designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{{\rm obs,} v}, \,\, \vt_{{\rm obs,} f}]$.  For 
the former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented here for clarity.}
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite. Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000. Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$. Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
\label{fig:broadening}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $\vt_{{\rm obs}, f} = 
[\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it {\it internal} to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-order (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of 
how these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
spectrum (of $N_{\rm pix}$ pixels) can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\vt, \vT_{\rm cheb}) \equiv \vD-\vM(\vt, \vT_{\rm cheb}).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:prob}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnprob}
\end{equation}
